---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
options(width = 108)
```

```{r include=FALSE}
library(dplyr)
library(magrittr)
library(purrr)
library(rlang)
library(lambdass)
```


# lambda syntax-sugar (lambdass)

## What is this package?
The purpose of this package is to provide you with an easy syntax for making an anonymous
function. 

## Related packages and functions
* `rlang::new_function()`, `rlang::as_closure()` https://github.com/tidyverse/rlang
* `pryr::f()`, `pryr::make_function()` https://github.com/hadley/pryr
* `lambda.r::"%as%"`  https://github.com/zatonovo/lambda.r
* `lambdaR::lambda()`  https://github.com/hoxo-m/lambdaR    

## Similar notation
Family of tidyverse functions such as `dplyr::mutate_if` and `purrr::map` 
whose formal arguments include `.predicate`, `.f`, and/or `.funs`
for a closure object can accept formula 
notation with `..1` or `.x` which is converted to a closure by `rlang::as_closure()`.
See the comparison of usage below.

```{r}
if ("package:lambdass" %in% search()) detach("package:lambdass")
library(purrr)
microbenchmark::microbenchmark(
  1 %>% map(~ ..1 + 1),    # purrr::map with rlang:::as_closure
  1 %>% map(~ .x + 1),     # purrr::map with rlang:::as_closure
  times = 1e3
)

suppressPackageStartupMessages(library("lambdass"))
microbenchmark::microbenchmark(
  1 %>% Map(~~ .. + 1, .), # base::Map with lamdass's double-tilde
  1 %>% map(~~ .. + 1),    # purrr::map with lambdass's double-tilde
  1 %>% map(~~ ..1 + 1),   # purrr::map with lambdass's double-tilde
  times = 1e3
)
```

## Installation
```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("tobcap/lambdass")
library("lambdass")
```

## Usage
can be written in three ways.
```{r eval=FALSE}
~~ .. + 1
f.(x, x + 1)
x %->% {x + 1}

# all means `add one`
function(x) x + 1
```

### Double-tilde
Double-tilde with dotted symbol placeholder makes an anonymous function
like the usage of `%` in Closure's lambda, `_` in Scala's lambda, or `#`
in Mathematica's Pure Function. See details in the documents below.

- http://en.wikibooks.org/wiki/Clojure_Programming/Examples/API_Examples/Function_Tools#.25
- http://www.scala-lang.org/files/archive/spec/2.11/06-expressions.html#placeholder-syntax-for-anonymous-functions
- http://reference.wolfram.com/language/howto/WorkWithPureFunctions.html  

A bounded variable can be specified by `..` which is a synonym for `..1`.
Arguments can be designated by `..1`, `..2`, up to `..5`.

```{r eval=FALSE}
~~ .. + 1
~~ ..2 / ..1

Map(~~ .. ^ 2, 1:10)
Reduce(~~ ..1 + ..2, 1:10)
```

The placeholder must be in order.
```{r error=TRUE}
~~ ..2 + 1
```

Double-tilde cannot create curried-function such as `function(x) function(y) x + y`.
The "De Bruijn index" is suitable for a notation of lambda calculas, 
but double-tilde is designed to cooperate with normal use of R's functions,
not curried-function.
See https://en.wikipedia.org/wiki/De_Bruijn_index for "De Bruijn index".


### f.
A function name `f()` is used in many situations, so I avoid using it and `f.()` 
is adopted.

```{r eval=FALSE}
# f.(...arguments, body)
f.(x, x + 1)
f.(x, y, x + y)
# can define curried-function
f.(x, f.(y, x + y)) 

Map(f.(x, x ^ 2), 1:10)
Reduce(f.(x, y, x + y), 1:10)
```

`f.()` is now implemented by C.
The conceptual R code is defined by `f.r()`.
```{r}
identical(f.(x, y, x + y), f.r(x, y, x + y))

## about ten times speed-up
microbenchmark::microbenchmark(
  f.(x, y, x + y), 
  f.r(x, y, x + y)
)
```

### Arrow-notation
right-hand side always needs `{` because R cannot control the
strength of associativity for `%infix-function%`.

```{r eval=FALSE}
x %->% {x + 1}

# left-hand side is written in two ways:
{x; y} %->% {x + y}
`any-valid-varname`(x, y) %->% {x + y}
# but recommend to use just `f`, for readability and simplicity
f(x, y) %->% {x + y}

Map(x %->% {x ^ 2}, 1:10)
Reduce({x; y} %->% {x + y}, 1:10)
```

Arguments can be checked when using `:` with the result of
`tyepof(actual_argument)` as symbol
```{r}
{x:integer} %->% {x * 2}
f(x:character) %->% {paste(x, "test")}
```

When the default value is set to a parameter, its type-checking 
syntax is automatically added at head of function-body.
```{r}
{x = 1L} %->% {x * 2}
f(x:character, y = "") %->% {paste0(x, y)}
```

## Benchmarking
Note that it's nanoseconds
```{r eval=FALSE}
microbenchmark::microbenchmark(
   "f." = f.(x, x), 
   "~~" = ~~ ..,
   "%->%" = {x} %->% {x},
   "as.function" = as.function(alist(x=, x)),
   "as.function.default" = as.function.default(alist(x=, x)),
   "function(x) x" = function(x) x
)
# Unit: nanoseconds
#                 expr    min       lq      mean   median     uq    max neval
#                   f.   5350   6687.5   8073.96   8025.0   8916  34326   100
#                   ~~  10699  12037.0  13704.19  13374.0  14712  28977   100
#                 %->% 138194 141983.5 160380.99 145772.5 170514 324978   100
#          as.function  32543  33881.0  41360.74  36110.0  40567 188122   100
#  as.function.default  22736  24295.5  27447.78  26302.0  29423  66869   100
#        function(x) x      0    446.0    910.25    447.0    892  38338   100

microbenchmark::microbenchmark(
   "f." = f.(x, y, x + y), 
   "~~" =  ~~ ..1 + ..2,
   "%->%" = f(x, y) %->% {x + y},
   "as.function" = as.function(alist(x=, y=, x + y)),
   "as.function.default" = as.function.default(alist(x=, y=, x + y)),
   "function(x, y) x+y" = function(x, y) x + y
)
# Unit: nanoseconds
#                 expr    min     lq      mean   median       uq    max neval
#                   f.   6241   7134   8947.70   8471.0   9362.5  20507   100
#                   ~~  13374  14712  19263.05  18724.0  20506.0  98965   100
#                 %->% 113676 119916 131948.24 122814.5 128163.0 327652   100
#          as.function  33434  35663  41761.75  39006.5  44133.0  74001   100
#  as.function.default  23627  25411  30964.91  27639.0  32766.0  73109   100
#   function(x, y) x+y      0    446    602.64    447.0    892.0   4904   100
```

Comparison with `rlang::as_closure()`
```{r}
microbenchmark::microbenchmark(
  ~~ ..,
  ~~ ..1,
  as_closure(~ .x),
  as_closure(~ ..1),
  identity
)
```

Apply closure
```{r}
microbenchmark::microbenchmark(
  (~~ ..)(1) ,
  (~~ ..1)(1) ,
  as_closure(~ .x)(1),
  as_closure(~ ..1)(1),
  identity(1)
)
```


## Church Encoding
By using f.(), R notation is almost as-is with the article written in Wikipedia.

https://github.com/TobCap/lambdass/blob/master/vignettes/ChurchEncoding.md
